---
title: "Statistical Graphics and Visualization"
description: |
  Companion website for 35-315 offered in Spring 2023 and taught by Professor Ron Yurko.
site: distill::distill_website
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)

# Learn more about creating websites with Distill at:
# https://rstudio.github.io/distill/website.html

# Learn more about publishing to GitHub Pages at:
# https://rstudio.github.io/distill/publish_website.html#github-pages

```


## Lectures

Lecture slides are available below in html format (see Canvas for PDF format), and the .Rmd files are linked from GitHub.

| Lecture | Topic                         | html                                                                               | Rmd                                                                                                              |
|---------|-----------------------------------------|-----------|-----------|
| 1       | Introduction and the Grammar of Graphics | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec1#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec1.Rmd) |
| 2       | 1D Categorical Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec2#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec2.Rmd) |
| 3       | Statistical Inference for 1D Categorical Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec3#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec3.Rmd) |
| 4       | Visualizing 1D Inference and 2D Categorical Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec4#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec4.Rmd) |
| 5       | Visualizing Inference for 2D Categorical Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec5#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec5.Rmd) |
| 6       | 1D Quantitative Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec6#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec6.Rmd) |
| 7       | Density Estimation | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec7#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec7.Rmd) |
| 8       | Graphical Inference for 1D Quantitative Data | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec8#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec8.Rmd) |
| 9       | Comparing Distributions and Statistical Power | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec9#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec9.Rmd) |
| 10      | 2D Quantitative Data: Scatterplots and Linear Regression | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec10#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec10.Rmd) |
| 11      | Inference with Linear Regression | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec11#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec11.Rmd) |
| 12      | More Regression and Midsemester Review | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec12#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec12.Rmd) |
| 13      | Nonlinear Regression and Pairs Plots | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/slides/Lec13#1) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/slides/Lec13.Rmd) |



## R Demos

Additional `R` demos are available below in html format.

| Demo  | Topic                | html   | Rmd       |
|-------|--------------------|--------|-----------|
| 1     | Into the tidyverse | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/demos/Demo1.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/demos/Demo1.Rmd) |
| 2     | Lecture 10 | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/demos/Lecture10_R_Demo.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/demos/Lecture10_R_Demo.Rmd) |
| 3     | Lecture 11 | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/demos/Lecture11_R_Demo.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/demos/Lecture11_R_Demo.Rmd) |
| 4     | Lecture 13 | [`r icons::fontawesome$solid$laptop`](https://cmu-36315-spring23.netlify.app/demos/Lecture13_R_Demo.html) | [`r icons::fontawesome$brands$github`](https://github.com/ryurko/Spring23-36315-Website/blob/main/demos/Lecture13_R_Demo.Rmd) |

## Resources

### Recommended textbooks

-   R for Data Science: <https://r4ds.had.co.nz/>

-   Fundamentals of Data Visualization: <https://clauswilke.com/dataviz/index.html>

-   Data Visualization: A Practical Introduction: <https://socviz.co/index.html#preface>

-   ggplot2: Elegant Graphics for Data Analysis: <https://ggplot2-book.org/>

-   Text Mining with R: <https://www.tidytextmining.com/>

### More `ggplot2` resources

-   The `ggplot2` cheatsheet: <https://www.maths.usyd.edu.au/u/UG/SM/STAT3022/r/current/Misc/data-visualization-2.1.pdf>

-   Creating custom themes: <https://themockup.blog/posts/2020-12-26-creating-and-using-custom-ggplot2-themes/>

-   R Graphics cookbook: <https://r-graphics.org/>

-   A guide to various scales: <https://ggplot2tor.com/scales>

### R Markdown

-   The definitive guide to R Markdown: <https://bookdown.org/yihui/rmarkdown/>

-   R Markdown reference guide: <https://www.rstudio.com/wp-content/uploads/2015/03/rmarkdown-reference.pdf>

-   R Markdown cheat sheet: <https://www.rstudio.com/wp-content/uploads/2015/02/rmarkdown-cheatsheet.pdf>

## Acknowledgments {.appendix}

This course website is made with the [distill package](https://rstudio.github.io/distill/). Thanks to [Meghan Hall](https://meghan.rbind.io/), [Alison Hill](https://rstudio-education.github.io/teaching-in-production/), and [Shamindra Shrotriya](https://www.shamindras.com/posts/2019-07-11-shrotriya2019distillpt1/) for their incredible examples and tutorials that helped in the process of making this site.

```{r, include=FALSE, results='asis'}
library(metathis)
meta() %>%
  meta_social(
    title = "CMU 36-315",
    description = "Course website for Statistical Graphics and Visualization, by Ron Yurko",
    url = "https://cmu-36315-spring23.netlify.app/",
    image_alt = "CMU 36-315",
    og_type = "website",
    og_author = "Ron Yurko",
    twitter_card_type = "summary",
    twitter_creator = "@Stat_Ron"
  )
```
